\tool: Programmable, Efficient, and Scalable Secure Two-Party Computation

### 2017-10-31 OaklandSubmission/eval.tex
para heads

\subsubsection*{Deep neural nets}
The next class of classifiers that we benchmark are deep neural nets
or DNNs. A DNN has multiple layers such that each layer computes a
matrix
multiplication followed by an {\it activation} function $f$. The most
common activation functions are square $f(x)=x^2$ and rectifier linear
unit (ReLU) $f(x)=\mathsf{max}(x,0)$.
If $A$ is a matrix then the output of $f(A)$ is also a matrix that is
obtained by applying $f$ to each entry of $A$ pointwise.
Given an input vector $x$, the predicted label of a DNN is
\[
 \mathsf{argmax}\ W_N\matmul f_{N-1}(\ldots f_1(W_1\matmul x)\ldots)
\]
Here, $f_i$'s are the (public) activation functions, the model
consists of matrices $W_i$,  $x \in \R^d$ is the input vector, and the
operator $\matmul$ denotes a matrix multiplication.
Neural nets usually have one or more fully connected layers, each of
which multiplies a matrix with a vector.
Some neural nets have convolution layers and such DNNs are also called
Convolutional Neural Nets or CNNs.
For the purpose of this paper, a convolution can be considered as a
(heavy) matrix-matrix multiplication. The size of matrices manipulated
by a convolution layer grows linearly with {\it window size}
(typically 9 or 25), the number of {\it output channels} (typically
16, 32, or 64), and the size of the matrix input to this layer.
Therefore, fully connected layers are lighter computation-wise compared
to convolution layers. However, the model size  of fully connected
layers is larger than those of convolution layers.
In general, DNNs are computationaly heavy but provide
much better accuracies on computer vision tasks than the classifiers
discussed above.

@@

In the following, we implement these classifiers in \tool and report
the time taken for making secure predictions. Ideally, the machine
learning classifiers are mathematical expressions over $\R$ that are
usually approximated by floating-point operations. 
As is standard, we port the classifers to integer manipulating
programs by scaling the models and rounding~\cite{minionn}. These
ported classifiers are then implemented in \tool.

@@

In Table~\ref{tab:nn}, 
the column ``Model size'' is the number of parameters in the
trained model.
We observe that our performance is competitive with specialized
\minion protocols, for both the LAN and the WAN settings. Further, lines
of \tool source code required is still small. 
We note that \minion and \tool implementations are both based on ABY.
These results are surprising as \minion has a much more efficient
preprocessing phase and has SIMD 
(single instruction multiple data) capabilities that allow it to perform
matrix operations efficiently. In contrast, \tool focuses on generic
2PC, and such optimizations are part of our future work.
\minion also reports performance results on a bigger DNN with 7
convolution layers. In \tool, this benchmark requires partitioning and
we discuss it in Section~\ref{sec:pipeeval}.
\\\\
\noindent\textbf{State-of-the-art classifiers.}
Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
Its introductory tutorial describes two prediction models for handwritten digit recognition
using the MNIST dataset~\cite{mnist}.
Each image in this dataset is a greyscale $28\times 28$ image of
digits 0 to 9.
The first model that the tutorial describes is a softmax regression
that provides an accuracy of 92\%. The classifier evaluates
$ \mathsf{argmax}\ W\matmul x+b$.
Here, $x$ is a 784 length vector obtained from the input image,
each $W$ is a $10\times 784$ matrix, and each $b$ is a $10$ length
vector.
We implement this classifer in \tool and present the results in the
first row of Table~\ref{tab:tf}. We note that a pure boolean circuit for this simple functionality does not fit in the memory of our machines.

The next classifer in the Tensorflow tutorial is a convolution neural net with two convolutions
(with 32 output channels) and two fully connected layers with ReLU as the activation function.
This DNN is both bigger and more accurate than the DNNs presented in the previous section.
In particular, it has an accuracy of 99.2\%.
Since, we are not aware of any other tools that have used this
model as a benchmark, we only report numbers for \tool.
We observe that this DNN can take a minute per prediction in the WAN
setting and is the largest benchmark that we have evaluated without
partitioning.

@@

We next present \bonsai~\cite{bonsai} results on three
standard datasets: character recognition (Chars4k~\cite{campos},
accuracy 74.71\%), text recognition (USPS~\cite{hull}, accuracy
94.4\%), and object categorization (WARD~\cite{yang}, accuracy
95.7\%). 
The \bonsai models are already over integers and no porting from
floating-point is required.
We implement the trained classifiers in \tool for all the benchmarks
from~\cite{bonsai},
and show the represenative results in Table~\ref{tab:bonsai}.
Out of all the benchmarks from~\cite{bonsai}, the dataset WARD
requires the largest model. 
The column ``depth''  shows the depth of the
tree used by \bonsai. The size of \tool 
program grows with the depth of the tree, as the straightforward \tool
implementation requires a loop for each layer of the
tree\footnote{We remark here that our current language does not support functions (which we leave for future work) and with this support, LOC would be lower and independent of depth.}.

@@

\setlength\tabcolsep{6pt}
\subsection{Secure code partitioning}
\label{sec:pipeeval}
\noindent\textbf{CIFAR-10}
The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
dataset~\cite{cifar}. 
The classifier's task is to categorize colored ($32\times 32$) images
into 10 classes. A secure evaluation of this DNN needs more memory
than what is available
on our machines. Therefore, we use partitioning and divide the
computation into seven stages. 
The first step does a convolution with 64 output channels and a ReLU
activation. The next four stages together perform a convolution that
involves multiplying a $64\times 576$ matrix with a 
$576\times 1024$ matrix. The sixth stage performs a ReLU and a
convolution. The final stage has four convolutions, five ReLUs, and a
fully connected layer. The total number of lines of \tool code for
this benchmark is 336 lines.

@@

 \caption{Partitioning results for matrix factorization. The time reported by~\cite{valeriaMatrix} for this computation is 10440 seconds.}
 \label{tab:factor} 
\end{table}
Table~\ref{tab:cifar} shows the end-to-end numbers as
well as the numbers for the sixth stage, which is the heaviest. 
The number of gates are in millions, hence the suffix `m' in the last
three columns.
%For this DNN, the specialized \minion protocol takes 544 seconds and
%communicates 9272 Mb.
As with Table~\ref{tab:nn}, \tool generated generic 2PC protocol is
competitive with \minion here as well. Therefore,
we believe that with partitioning, \tool can scale to arbitrary sized
computations while maintaining performance
competitive with existing specialized protocols. In particular, for a
large enough DNN, \minion could run
out of memory but an appropriately partitioned \tool implementation would still
succeed.
\\\\
\noindent\textbf{Matrix factorization}
\tool is not tied to secure prediction and can express more general computations.
To demonstrate this expressiveness, we  implement secure matrix factorization~\cite{valeriaMatrix}. Abstractly, given a sparse matrix $\mathcal{M}$ of dimensions
$n\times m$ and $M$ non-zero entries, the goal is to generate a matrix $U$ of dimension $n\times d$ and a matrix
$V$ of dimension $d\times m$ such that $\mathcal{M}\approx UV$. This operator is useful in recommender systems.
In particular, Nikolaenko et al.~\cite{valeriaMatrix} shows how to implement a movie recommender system which does not require users to reveal their data in the clear, i.e., the ratings the users have assigned to movies are kept secret. The implementation is a two party computation of an iterative algorithm for matrix factorization (Algorithm~1 in~\cite{valeriaMatrix}).
This algorithm is based on gradient descent and iteratively converges to a local minima.
We implement this algorithm in \tool.

To ensure that the algorithm converges to the right local minima, Nikolaenko et al. require
36 bits of precision. Since ABY supports either 32-bit or 64-bit integers, our \tool implementation
manipulates 64-bit variables. For the matrix $\mathcal{M}$ of  user data, Nikolaenko et al. consider $n=940$ users, $m=40$ most popular movies, and $M=14683$ ratings from the MovieLens dataset. The time reported in~\cite{valeriaMatrix}
for one iteration is 2.9 hours\footnote{~\cite{valeriaMatrix} does not report the network round-trip time.}. This compuation is large enough that we partition each iteration
into three stages. The first stage involves a Batcher~\cite{Batcher} sorting network followed by a linear pass.
The second stage involves sorting and gradient computations and is the heaviest stage.
The third  stage is similar to the first stage. The results are presented in Table~\ref{tab:factor}.

>>>>

\subsubsection*{Deep neural nets}
The next class of classifiers that we benchmark are deep neural nets
or DNNs. A DNN has multiple layers such that each layer computes a
matrix
multiplication followed by an {\it activation} function $f$. The most
common activation functions are square $f(x)=x^2$ and rectifier linear
unit (ReLU) $f(x)=\mathsf{max}(x,0)$.
If $A$ is a matrix then the output of $f(A)$ is also a matrix that is
obtained by applying $f$ to each entry of $A$ pointwise.
Given an input vector $x$, the predicted label of a DNN is
\[
 \mathsf{argmax}\ W_N\matmul f_{N-1}(\ldots f_1(W_1\matmul x)\ldots)
\]
Here, $f_i$'s are the (public) activation functions, the model
consists of matrices $W_i$,  $x \in \R^d$ is the input vector, and the
operator $\matmul$ denotes a matrix multiplication.
Neural nets usually have one or more fully connected layers, each of
which multiplies a matrix with a vector.
Some neural nets have convolution layers and such DNNs are also called
Convolutional Neural Nets or CNNs.
For the purpose of this paper, a convolution can be considered as a
(heavy) matrix-matrix multiplication. The size of matrices manipulated
by a convolution layer grows linearly with {\it window size}
(typically 9 or 25), the number of {\it output channels} (typically
16, 32, or 64), and the size of the matrix input to this layer.
Therefore, fully connected layers are lighter computation-wise compared
to convolution layers. However, the model size  of fully connected
layers is larger than those of convolution layers.
In general, DNNs are computationally heavy but provide
much better accuracies on computer vision tasks than the classifiers
discussed above.

@@

In the following, we implement these classifiers in \tool and report
the time taken for making secure predictions. Ideally, the machine
learning classifiers are mathematical expressions over $\R$ that are
usually approximated by floating-point operations. 
As is standard, we port the classifiers to integer manipulating
programs by scaling the models and rounding~\cite{minionn}. These
ported classifiers are then implemented in \tool.

@@

In Table~\ref{tab:nn}, 
the column ``Model size'' is the number of parameters in the
trained model.
We observe that our performance is competitive with specialized
\minion protocols, for both the LAN and the WAN settings. Further, lines
of \tool source code required is still small. 
We note that \minion and \tool implementations are both based on ABY.
These results are surprising as \minion has a much more efficient
preprocessing phase and has SIMD 
(single instruction multiple data) capabilities that allow it to perform
matrix operations efficiently. In contrast, \tool focuses on generic
2PC, and such optimizations are part of our future work.
\minion also reports performance results on a bigger DNN with 7
convolution layers. In \tool, this benchmark requires partitioning and
we discuss it in Section~\ref{sec:pipeeval}.
\\\\
\noindent\textbf{State-of-the-art classifiers.}
Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
Its introductory tutorial describes two prediction models for handwritten digit recognition
using the MNIST dataset~\cite{mnist}.
Each image in this dataset is a greyscale $28\times 28$ image of
digits 0 to 9.
The first model that the tutorial describes is a softmax regression
that provides an accuracy of 92\%. The classifier evaluates
$ \mathsf{argmax}\ W\matmul x+b$.
Here, $x$ is a 784 length vector obtained from the input image,
each $W$ is a $10\times 784$ matrix, and each $b$ is a $10$ length
vector.
We implement this classifier in \tool and present the results in the
first row of Table~\ref{tab:tf}. We note that a pure boolean circuit for this simple functionality does not fit in the memory of our machines.

The next classifier in the Tensorflow tutorial is a convolution neural net with two convolutions
(with 32 output channels) and two fully connected layers with ReLU as the activation function.
This DNN is both bigger and more accurate than the DNNs presented in the previous section.
In particular, it has an accuracy of 99.2\%.
Since, we are not aware of any other tools that have used this
model as a benchmark, we only report numbers for \tool.
We observe that this DNN can take a minute per prediction in the WAN
setting and is the largest benchmark that we have evaluated without
partitioning.

@@

We next present \bonsai~\cite{bonsai} results on three
standard datasets: character recognition (Chars4k~\cite{campos},
accuracy 74.71\%), text recognition (USPS~\cite{hull}, accuracy
94.4\%), and object categorization (WARD~\cite{yang}, accuracy
95.7\%). 
The \bonsai models are already over integers and no porting from
floating-point is required.
We implement the trained classifiers in \tool for all the benchmarks
from~\cite{bonsai},
and show the representative results in Table~\ref{tab:bonsai}.
Out of all the benchmarks from~\cite{bonsai}, the dataset WARD
requires the largest model. 
The column ``depth''  shows the depth of the
tree used by \bonsai. The size of \tool 
program grows with the depth of the tree, as the straightforward \tool
implementation requires a loop for each layer of the
tree\footnote{We remark here that our current language does not support functions (which we leave for future work) and with this support, LOC would be lower and independent of depth.}.

@@

\setlength\tabcolsep{6pt}
\subsection{Secure code partitioning}
\label{sec:pipeeval}
\noindent\textbf{CIFAR-10.}
The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
dataset~\cite{cifar}. 
The classifier's task is to categorize colored ($32\times 32$) images
into 10 classes. A secure evaluation of this DNN needs more memory
than what is available
on our machines. Therefore, we use partitioning and divide the
computation into seven stages. 
The first step does a convolution with 64 output channels and a ReLU
activation. The next four stages together perform a convolution that
involves multiplying a $64\times 576$ matrix with a 
$576\times 1024$ matrix. The sixth stage performs a ReLU and a
convolution. The final stage has four convolutions, five ReLUs, and a
fully connected layer. The total number of lines of \tool code for
this benchmark is 336 lines.

@@

 \caption{Partitioning results for matrix factorization. The time reported by~\cite{valeriaMatrix} for this computation is 10440 seconds.}
 \label{tab:factor} 
\end{table}
Table~\ref{tab:cifar} shows the end-to-end numbers as
well as the numbers for the sixth stage, which is the heaviest. 
The number of gates are in millions, hence the suffix `m' in the last
three columns.
%For this DNN, the specialized \minion protocol takes 544 seconds and
%communicates 9272 Mb.
As with Table~\ref{tab:nn}, \tool generated generic 2PC protocol is
competitive with \minion here as well. Therefore,
we believe that with partitioning, \tool can scale to arbitrary sized
computations while maintaining performance
competitive with existing specialized protocols. In particular, for a
large enough DNN, \minion could run
out of memory but an appropriately partitioned \tool implementation would still
succeed.
\\\\
\noindent\textbf{Matrix factorization.}
\tool is not tied to secure prediction and can express more general computations.
To demonstrate this expressiveness, we  implement secure matrix factorization~\cite{valeriaMatrix}. Abstractly, given a sparse matrix $\mathcal{M}$ of dimensions
$n\times m$ and $M$ non-zero entries, the goal is to generate a matrix $U$ of dimension $n\times d$ and a matrix
$V$ of dimension $d\times m$ such that $\mathcal{M}\approx UV$. This operator is useful in recommender systems.
In particular, Nikolaenko et al.~\cite{valeriaMatrix} shows how to implement a movie recommender system which does not require users to reveal their data in the clear, i.e., the ratings the users have assigned to movies are kept secret. The implementation is a two party computation of an iterative algorithm for matrix factorization (Algorithm~1 in~\cite{valeriaMatrix}).
This algorithm is based on gradient descent and iteratively converges to a local minima.
We implement this algorithm in \tool.

To ensure that the algorithm converges to the right local minima, Nikolaenko et al. require
36 bits of precision. Since ABY supports either 32-bit or 64-bit integers, our \tool implementation
manipulates 64-bit variables. For the matrix $\mathcal{M}$ of  user data, Nikolaenko et al. consider $n=940$ users, $m=40$ most popular movies, and $M=14683$ ratings from the MovieLens dataset. The time reported in~\cite{valeriaMatrix}
for one iteration is 2.9 hours\footnote{~\cite{valeriaMatrix} does not report the network round-trip time.}. This computation is large enough that we partition each iteration
into three stages. The first stage involves a Batcher~\cite{Batcher} sorting network followed by a linear pass.
The second stage involves sorting and gradient computations and is the heaviest stage.
The third  stage is similar to the first stage. The results are presented in Table~\ref{tab:factor}.
